### 2012-05-23 klein2012/headertikz.tex
Déplacement des slides dans le bon dossier, completion des slides et correction des notations

\usepackage{pgf}
\usepackage{pgfcore}
\usepackage{pgfbaseimage}
\usepackage{pgfbaselayers}
\usepackage{pgfbasepatterns}
\usepackage{pgfbaseplot}
\usepackage{pgfbaseshapes}
\usepackage{pgfbasesnakes}
\usepackage{tikz}

\tikzset{normal/.style={ 
% The shape: 
rectangle,minimum size=6mm,rounded corners=3mm, 
% The rest 
very thick,draw=black!50, 
top color=white,bottom color=black!20, 
font=\ttfamily}}
\tikzset{arduino/.style={ 
% The shape: 
rectangle, 
% The size: 
minimum size=6mm, 
% The border: 
very thick, 
draw=blue!50!black!50, % 50% red and 50% black, 
% and that mixed with 50% white 
% The filling: 
top color=white, % a shading that is white at the top... 
bottom color=blue!50!black!20, % and something else at the bottom 
% Font 
font=\itshape 
}}
\tikzset{enceinte/.style={ 
% The shape: 
rectangle, 
% The size: 
minimum size=6mm, 
% The border: 
very thick, 
draw=yellow!90!black!50, % 50% red and 50% black, 
% and that mixed with 50% white 
% The filling: 
top color=white, % a shading that is white at the top... 
bottom color=yellow!90!black!20, % and something else at the bottom 
% Font 
font=\itshape 
}}

\tikzset{ordi/.style={ 
% The shape: 
rectangle, 
% The size: 
minimum size=6mm, 
% The border: 
very thick, 
draw=green!90!black!50, % 50% red and 50% black, 
% and that mixed with 50% white 
% The filling: 
top color=white, % a shading that is white at the top... 
bottom color=green!90!black!20, % and something else at the bottom 
% Font 
font=\itshape 
}}